Draft Exploration

### 2020-04-06 draft_exploration.Rmd
made mlb percent functions for signed and unsigned high schoolers

@@

```{r, echo=FALSE}
redraft_table %>%
  ggplot(aes(round, mean_round_diff)) +
  geom_point(na.rm = TRUE) +
  geom_line(aes(group = start_year, color = start_year), na.rm = TRUE) +
  labs(
    x = "Round Range Drafted out of High School",
    y = "Average Improvement Re-draft Round",
    color = "Time Frame",
    title = "Average Change in Round of Re-drafted Unsigned High School Picks",
    subtitle = "Organized by initial round range and time period"
  ) +
  theme_light() +
  scale_color_viridis_c() +
  scale_x_continuous(breaks = c(5, 10, 20, 30, 40, 50)) +
  scale_y_continuous(breaks = c(-10, -5, 0, 10, 20)) +
  geom_hline(color = "red", linetype = "dashed", yintercept = 0)
```

@@

>>>>

```
Notes on graph: most recent 5 years emulates earliest two 5 year periods but much less steep towards back end of draft. Possible explanations are teams taking late draft shots at unsignable top picks or high school picks just maintaining stock better.

@@

```{r, echo=FALSE}
redraft_table %>%
  ggplot(aes(round, mean_round_diff)) +
  geom_point(na.rm = TRUE) +
  geom_line(aes(group = start_year, color = start_year), na.rm = TRUE) +
  labs(
    x = "Round Range Drafted out of High School",
    y = "Average Improvement Re-draft Round",
    color = "Time Frame",
    title = "Average Change in Round of Re-drafted Unsigned High School Picks",
    subtitle = "Organized by initial round range and time period"
  ) +
  theme_light() +
  scale_color_viridis_c() +
  scale_x_continuous(breaks = c(5, 10, 20, 30, 40, 50)) +
  scale_y_continuous(breaks = c(-10, -5, 0, 10, 20)) +
  geom_hline(color = "red", linetype = "dashed", yintercept = 0)
```
Note estimated intercept with y=0 to see when on average it becomes beneficial
to a player's redraft stock to not sign out of high school. Important to keep in
mind this is for players who are redrafted, so back half of draft improvement
likely partly explained by fewer, the best, initial late round high school picks
being redrafted.

@@

mlb_percentile_signed <- function(round = 20, startyear = 1990, endyear = 2010) {

top_20_signed_total <- master %>%
  filter(source == "H" & signed == "Y") %>%
  filter(draft_round <= round & year >= startyear & year <= endyear) %>%
  count() %>%
  pull(n)

top_20_signed_mlb <- master %>%
  filter(source == "H" & signed == "Y" & high_level == "MLB") %>%
  filter(draft_round <= round & year >= startyear & year <= endyear) %>%
  count() %>%
  pull(n)

top_20_signed_mlb / top_20_signed_total

# 27.3% of signed top 20 round high school picks from 1990-2010 have made it to MLB

```{r, echo=FALSE}
top_20_unsigned_total <- master %>%
  filter(source == "H" & signed == "N" & draft_numb == 1) %>%
  filter(draft_round <= 20 & year >= 1990 & year <= 2010) %>%
  count() %>%
  pull(n)

top_20_unsigned_mlb <- master %>%
  filter(source == "H" & signed == "N" & draft_numb == 1 & high_level == "MLB") %>%
  filter(draft_round <= 20 & year >= 1990 & year <= 2010) %>%
  count() %>%
  pull(n)

mlb_percentile_20u <- top_20_unsigned_mlb / top_20_unsigned_total

# 22.7% of unsigned top 20 round high school picks from 1990-2010 have made it to MLB
